Publications included in this thesis:
\begin{enumerate}
    \item \textbf{Larsen S. J.}, Baumbach J. (2017). CytoMCS: a multiple maximum common subgraph detection tool for Cytoscape. In \emph{Proceedings of the 2017 International Symposium on Integrative Bioinformatics}.
    \item \textbf{Larsen S. J.}, do Canto L. M., Rogatto S. R., Baumbach J. (2018). \mbox{CoNVaQ}: a web tool for copy number variation-based association studies. \emph{BMC Genomics}. Springer Nature.
	\item \textbf{Larsen S. J.}, R\"ottger R., Schmidt H. H. H. W., Baumbach J. (2018). E. coli gene regulatory networks are inconsistent with gene expression data. \emph{Nucleic Acids Research}. Oxford University press.
    \item \textbf{Larsen S. J.}, Schmidt H. H. H. W., Baumbach J. (2019). \emph{De novo} and supervised endophenotyping using network-guided ensemble learning. (Under review)
    \item \textbf{Larsen S. J.}, Carlin D. E., Ideker T., Baumbach J. (2019). Analysis of genetic variants in chronic obstructive pulmonary disease using a hierarchical model of the cell. (Unpublished)
\end{enumerate}
Other contributions not included in this thesis:
\begin{itemize}
    \item Costa A. A. B. A., do Canto L. M., \textbf{Larsen S. J.}, Ribeiro A. R. G., Stecca C. E., Petersen A. H., Aagaard M. M., de Brot L., Baumbach J., Baiocchi G., Achatz M. I., Rogatto S. R. (2019). Genomic profiling in ovarian cancer retreated with platinum based chemotherapy presented homologous recombination deficiency and copy number imbalances of CCNE1 and RB1 genes. \emph{BMC Cancer}. Springer Nature.
    \item do Canto L. M., \textbf{Larsen S. J.}, Kupper B. E. C., de Souza Begnami M. D. F., Scapulatempo-Neto C., Petersen A. H., Aagaard M. M., Baumbach J., Aguiar-Jr S., Rogatto S. R. (2019). Increased levels of genomic instability and mutations in homologous recombination genes in locally advanced rectal carcinomas. \emph{Frontiers in Oncology}. Frontiers.
    \item Laufer-Amorim R., Fonseca-Alves C. E., Villacis R. A. R., Linde S. A. D., Carvalho M., \textbf{Larsen S. J.}, Marchi F. A., Rogatto S. R. (2019). Comprehensive genomic profiling of androgen-receptor-negative canine prostate cancer. \emph{International Journal of Molecular Sciences}. MDPI.
    \item Casas A. I., Hassan A. A., \textbf{Larsen S. J.}, Gomez-Rangel V., Elbatreek M., Kleikers P. W. M., Guney E., Egea J., López M. G., Baumbach J., Schmidt H. H. H. W. (2019). From single drug targets to synergistic network pharmacology in ischemic stroke. \emph{Proceedings of the National Academy of Sciences}. National Academy of Sciences.
    \item Tweedell R. E., Tao D., Hamerly T., Robinson T. M., \textbf{Larsen S. J.}, Gr\o{}nning A. G., Norris A. M., King J. G., Law H. C., Baumbach J., Bergmann-Leitner E. S. (2019). The Selection of a Hepatocyte Cell Line Susceptible to Plasmodium falciparum Sporozoite Invasion That is Associated With Expression of Glypican-3. \emph{Frontiers in Microbiology}. Frontiers.
    \item Grønning A. G. B., Doktor T. K., \textbf{Larsen S. J.}, Petersen U. S. S., Holm L. L., Bruun G. H., Hansen M. B., Hartung A., Baumbach J., Andresen B. S. (2019). DeepCLIP: Predicting the effect of mutations on protein-RNA binding with Deep Learning. (Submitted)
\end{itemize}
